The Role of Shallow Processing in Agreement Attraction

### 2020-02-07 agremeent_draft_dec17/draft_dec17.Rnw
Added simpler by-experiment analyses.

@@

<<exp1ResponseModel, fig.cap='Estimates and 95\\% credible intervals for the regression coefficients for Experiment 1.', fig.height=3, fig.align='center'>>=
contrast_names <- c("cUngrammatical" = "Ungrammaticality",
                    "cAttractorPlural" = "Plural Attactor",
                    "cUngrammatical:cAttractorPlural" = "Ungrammaticality * Plural Attractor",
                    "cEndsInConsonant" = "Ambiguity",
                    "cEndsInConsonant:cUngrammatical" = "Ambiguity * Ungrammaticality",
                    "cEndsInConsonant:cAttractorPlural" = "Ambiguity * Plural Attractor",
                    "cEndsInConsonant:cUngrammatical:cAttractorPlural" = "Ambiguity * Ungrammaticality * Plural Attractor")

@@

@@

# avg_clean2 <- df_merged2 %>%
#                 group_by(experiment, source, grammatical, attractor_num) %>%
#                 summarize(avgRT = mean(RT), p_yes = mean(ResponseYes, na.rm = T))
# # XXX

@@

fname_exp2_responses <- "../workspace_exp2/fit_responses"

@@

# TODO: Data seems to contain NAs in relevant columns; fix it
m_responses2 <- brm(ResponseYes ~ cVerbalAttractor * cUngrammatical * cAttractorPlural + 
                                 (cUngrammatical * cAttractorPlural + 1| subject) +
                                 (cUngrammatical * cAttractorPlural + 1| item),
                   data = df_merged_nofillers2 %>% subset(!is.na(ResponseYes)),
                   family = bernoulli("probit"),
                   file = fname_exp2_responses, 
                   chains = n_chains, cores = n_cores, iter = n_iter)

@@

<<exp2ResponseModel, fig.cap='Estimates and 95\\% credible intervals for the regression coefficients for Experiment 2', fig.height=3, fig.align='center'>>=

@@

p_m_response_exp2 <- p_m_response_exp2 + annotate(x=-3, xend=1, y=0, yend=0, lwd=0.25, geom="segment")
print(p_m_response_exp2)

>>>>

fname_exp1_responses_exp1only <- "../workspace_exp1/fit_responses_exp1only"
m_responses_exp1only <- brm(ResponseYes ~ cUngrammatical * cAttractorPlural + 
                                 (cUngrammatical * cAttractorPlural + 1| subject) + 
                                 (cUngrammatical * cAttractorPlural + 1| item),
                   data = df_merged_nofillers %>% subset(experiment == "Experiment 1"),
                   family = bernoulli("probit"), 
                   chains = n_chains, cores = n_cores, iter = n_iter,
                   file = fname_exp1_responses_exp1only)

fname_exp1_responses_Lagoonly <- "../workspace_exp1/fit_responses_Lagoonly"
m_responses_Lagoonly <- brm(ResponseYes ~ cUngrammatical * cAttractorPlural + 
                                 (cUngrammatical * cAttractorPlural + 1| subject) + 
                                 (cUngrammatical * cAttractorPlural + 1| item),
                   data = df_merged_nofillers %>% subset(experiment != "Experiment 1"),
                   family = bernoulli("probit"), 
                   chains = n_chains, cores = n_cores, iter = n_iter,
                   file = fname_exp1_responses_Lagoonly)

@@

<<exp1ResponseModel, fig.cap='Estimates and 95\\% credible intervals for the regression coefficients for the model of Experiment 1 and Lago et al. (2018).', fig.height=3, fig.align='center'>>=
contrast_names <- c("cUngrammatical" = "Ungrammaticality",
                    "cAttractorPlural" = "Plural Attactor",
                    "cUngrammatical:cAttractorPlural" = "Ungrammaticality * Plural Attractor",
                    "cEndsInConsonant" = "Ambiguity",
                    "cEndsInConsonant:cUngrammatical" = "Ambiguity * Ungrammaticality",
                    "cEndsInConsonant:cAttractorPlural" = "Ambiguity * Plural Attractor",
                    "cEndsInConsonant:cUngrammatical:cAttractorPlural" = "Ambiguity * Ungrammaticality * Plural Attractor")

@@

<<exp1ResponseModelSepExp1, fig.cap='Estimates and 95\\% credible intervals for the regression coefficients for Experiment 1.', fig.height=1.5, fig.align='center'>>=
contrast_names_exp1 <- c("cUngrammatical" = "Experiment 1: Ungrammaticality",
                         "cAttractorPlural" = "Experiment 1: Plural Attactor",
                         "cUngrammatical:cAttractorPlural" = "Experiment 1: Ungrammaticality * Plural Attractor"
                        )

p_m_response_exp1 <- 
  create_model_coefs_plot( m_responses_exp1only, 
        plot_stats = T, map_names = contrast_names_exp1,
        expand_right = 2.5, expand_top = 2, x_stat_adjust = 1.1,
        x_breaks = c(-3,-2,-1, 0,1, 2) ) + 
        xlab("Estimate (probit)")

suppressWarnings({
print(p_m_response_exp1 + annotate(x=-3, xend=2, y=0, yend=0, lwd=0.25, geom="segment"))
})

contrast_names_lago <- c("cUngrammatical" = "Lago et al.: Ungrammaticality",
                         "cAttractorPlural" = "Lago et al.: Plural Attactor",
                         "cUngrammatical:cAttractorPlural" = "Lago et al.: Ungrammaticality * Plural Attractor"
                        )

<<exp1ResponseModelSepExp2, fig.cap='Estimates and 95\\% credible intervals for the regression coefficients for Lago et al. (2018).', fig.height=1.5, fig.align='center'>>=
p_m_response_lago <- 
  create_model_coefs_plot( m_responses_Lagoonly, 
        plot_stats = T, map_names = contrast_names_lago,
        expand_right = 2.5, expand_top = 2, x_stat_adjust = 1.1,
        x_breaks = c(-3,-2,-1, 0,1, 2) ) + 
        xlab("Estimate (probit)")

suppressWarnings({
print(p_m_response_lago + annotate(x=-3, xend=2, y=0, yend=0, lwd=0.25, geom="segment"))
})

@@

@@

@@

fname_exp2_responses <- "../workspace_exp2/fit_responses"
m_responses2 <- brm(ResponseYes ~ cVerbalAttractor * cUngrammatical * cAttractorPlural + 
                                 (cUngrammatical * cAttractorPlural + 1| subject) +
                                 (cUngrammatical * cAttractorPlural + 1| item),
                   data = df_merged_nofillers2 %>% subset(!is.na(ResponseYes)),
                   family = bernoulli("probit"),
                   file = fname_exp2_responses, 
                   chains = n_chains, cores = n_cores, iter = n_iter)

fname_exp2only_responses <- "../workspace_exp2/fit_responses_exp2only"
m_responses2_exp2only <- brm(ResponseYes ~ cUngrammatical * cAttractorPlural + 
                                 (cUngrammatical * cAttractorPlural + 1| subject) +
                                 (cUngrammatical * cAttractorPlural + 1| item),
                   data = df_merged_nofillers2 %>% 
                              subset(!is.na(ResponseYes)) %>% 
                              subset(experiment == "Experiment 2"),
                   family = bernoulli("probit"),
                   file = fname_exp2only_responses, 
                   chains = n_chains, cores = n_cores, iter = n_iter)

@@

<<exp2ResponseModel, fig.cap='Estimates and 95\\% credible intervals for the regression coefficients for a model of Experiment 1 and Experiment 2.', fig.height=3, fig.align='center'>>=

@@

p_m_response_exp2 <- p_m_response_exp2 + annotate(x=-3, xend=1, y=0, yend=0, lwd=0.25, geom="segment")
print(p_m_response_exp2)
@

<<exp2ResponseModelExp2Only, fig.cap='Estimates and 95\\% credible intervals for the regression coefficients for Experiment 2.', fig.height=1.5, fig.align='center'>>=

p_m_response_exp2_exp2only <- 
  create_model_coefs_plot( m_responses2_exp2only, 
        plot_stats = T, map_names = contrast_names2_le,
        expand_right = 4.5, expand_top = 1.5, x_stat_adjust = .75,
        x_breaks = c(-3,-2,-1, 0, 1) ) + 
        xlab("Estimate (probit)")

p_m_response_exp2only <- p_m_response_exp2_exp2only + annotate(x=-3, xend=1, y=0, yend=0, lwd=0.25, geom="segment")
print(p_m_response_exp2only)

### 2020-02-07 agremeent_draft_dec17/draft_dec17.tex
Added simpler by-experiment analyses.

A Bayesian GLM assuming a Bernoulli distribution with a probit-link function was fit to participants' `acceptable' responses. The model's estimates and 95\% credible intervals are shown in \autoref{fig:exp1ResponseModel}. 
%
%They confirm the observations we stated above: the positive interaction between sentence grammaticality and attractor number is in line with a larger effect of attractor number in ungrammatical sentences. 
%
The main effect of \textit{ungrammaticality} ($\hat{\beta}=-3.09;$ $CI=[-3.34; -2.85];$ $P(\beta<0)> .999$) indicates that, on average, participants were quite good at distinguishing between grammatical and ungrammatical sentences. Meanwhile, the positive interaction between \textit{ungrammaticality} and \textit{attractor number} ($\hat{\beta}=0.72;$ $CI=[0.41; 1.02];$ $P(\beta<0)< .001$) indicated a larger effect of attractor number in ungrammatical conditions, and thus a number agreement attraction effect.
Importantly, there was no evidence for a three-way interaction between \textit{the presence of ambiguity}, \textit{ungrammaticality} and \textit{attractor number} ($\hat{\beta}=0.31;$ $CI=[-0.21; 0.83];$ $P(\beta<0)=    .12$).

@@

\caption[Estimates and 95\% credible intervals for the regression coefficients for Experiment 1]{Estimates and 95\% credible intervals for the regression coefficients for Experiment 1.}\label{fig:exp1ResponseModel}
\end{figure}

@@

A Bayesian GLM assuming a Bernoulli distribution with a probit-link function was fit to participants' `acceptable' responses. The model's estimates and 95\% credible intervals are shown in \autoref{fig:exp2ResponseModel}. 
The main effect of \textit{ungrammaticality} ($\hat{\beta}=-3.41;$ $CI=[-3.63; -3.19];$ $P(\beta<0)> .999$) indicates that, 
on average, participants were quite good at distinguishing between grammatical and ungrammatical sentences. 
Meanwhile, there was no evidence for an interaction between \textit{ungrammaticality} and \textit{attractor number} ($\hat{\beta}=-0.03;$ $CI=[-0.33; 0.25];$ $P(\beta<0)=    .58$), indicating 
a similar effect of attractor number in ungrammatical conditions compared to grammatical conditions, and thus no number agreement attraction effect.
Importantly, the negative effect between \textit{the presence of the verbal plural}, \textit{ungrammaticality} and \textit{attractor number} ($\hat{\beta}=-1.07;$ $CI=[-1.53; -0.61];$ $P(\beta<0)> .999$) indicated that the plural morpheme on the verb of the object relative clause lowered the agreement attraction effects.

@@

\caption[Estimates and 95\% credible intervals for the regression coefficients for Experiment 2]{Estimates and 95\% credible intervals for the regression coefficients for Experiment 2}\label{fig:exp2ResponseModel}
\end{figure}

>>>>

A Bayesian GLM assuming a Bernoulli distribution with a probit-link function was fit to participants' `acceptable' responses. The model's estimates and 95\% credible intervals are shown in \autoref{fig:exp1ResponseModel}. 
%
%They confirm the observations we stated above: the positive interaction between sentence grammaticality and attractor number is in line with a larger effect of attractor number in ungrammatical sentences. 
%
The main effect of \textit{ungrammaticality} ($\hat{\beta}=-3.09;$ $CI=[-3.35; -2.84];$ $P(\beta<0)> .999$) indicates that, on average, participants were quite good at distinguishing between grammatical and ungrammatical sentences. Meanwhile, the positive interaction between \textit{ungrammaticality} and \textit{attractor number} ($\hat{\beta}=0.72;$ $CI=[0.42; 1.02];$ $P(\beta<0)< .001$) indicated a larger effect of attractor number in ungrammatical conditions, and thus a number agreement attraction effect.
Importantly, there was no evidence for a three-way interaction between \textit{the presence of ambiguity}, \textit{ungrammaticality} and \textit{attractor number} ($\hat{\beta}=0.30;$ $CI=[-0.20; 0.79];$ $P(\beta<0)=    .12$).

@@

\caption[Estimates and 95\% credible intervals for the regression coefficients for the model of Experiment 1 and Lago et al]{Estimates and 95\% credible intervals for the regression coefficients for the model of Experiment 1 and Lago et al. (2018).}\label{fig:exp1ResponseModel}
\end{figure}

{\centering \includegraphics[width=\maxwidth]{figure/exp1ResponseModelSepExp1-1}

\caption[Estimates and 95\% credible intervals for the regression coefficients for Experiment 1]{Estimates and 95\% credible intervals for the regression coefficients for Experiment 1.}\label{fig:exp1ResponseModelSepExp1}
\end{figure}

{\centering \includegraphics[width=\maxwidth]{figure/exp1ResponseModelSepExp2-1}

\caption[Estimates and 95\% credible intervals for the regression coefficients for Lago et al]{Estimates and 95\% credible intervals for the regression coefficients for Lago et al. (2018).}\label{fig:exp1ResponseModelSepExp2}
\end{figure}

@@

A Bayesian GLM assuming a Bernoulli distribution with a probit-link function was fit to participants' `acceptable' responses. The model's estimates and 95\% credible intervals are shown in \autoref{fig:exp2ResponseModel}. 
The main effect of \textit{ungrammaticality} ($\hat{\beta}=-3.09;$ $CI=[-3.31; -2.88];$ $P(\beta<0)> .999$) indicates that, 
on average, participants were quite good at distinguishing between grammatical and ungrammatical sentences. 
Meanwhile, there was no evidence for an interaction between \textit{ungrammaticality} and \textit{attractor number} ($\hat{\beta}=0.06;$ $CI=[-0.18; 0.30];$ $P(\beta<0)=    .30$), indicating 
a similar effect of attractor number in ungrammatical conditions compared to grammatical conditions, and thus no number agreement attraction effect.
Importantly, the negative effect between \textit{the presence of the verbal plural}, \textit{ungrammaticality} and \textit{attractor number} ($\hat{\beta}=-0.94;$ $CI=[-1.31; -0.57];$ $P(\beta<0)> .999$) indicated that the plural morpheme on the verb of the object relative clause lowered the agreement attraction effects.

@@

\caption[Estimates and 95\% credible intervals for the regression coefficients for a model of Experiment 1 and Experiment 2]{Estimates and 95\% credible intervals for the regression coefficients for a model of Experiment 1 and Experiment 2.}\label{fig:exp2ResponseModel}
\end{figure}

{\centering \includegraphics[width=\maxwidth]{figure/exp2ResponseModelExp2Only-1}

\caption[Estimates and 95\% credible intervals for the regression coefficients for Experiment 2]{Estimates and 95\% credible intervals for the regression coefficients for Experiment 2.}\label{fig:exp2ResponseModelExp2Only}
\end{figure}
